Asymmetric tail-association in extreme climatic events affects population-extinction risk

### 2019-10-11 MainText.Rmd
extreme and moderate results combined in single plot

***Shyamolina: species reordering is used to access the change in community 
in response to extreme events or global environmental changes, it relates to change 
in relative abundances of dominant species (mainly). For example, as different species have different 
sensitivity levels to extreme condition (means the extreme threshold is not 
constant for all the species in a community), some years with extreme drought favors say c3 grass 
but in the next year c4 grass could be dominant. Such alteration in community composition has "impact on 
ecosystem function, leading to an extreme ecological response". For detailed overview you may see 
@hoover2014resistance, Jones et al. Ecol. lett. 2017 (doi: 10.1111/ele.12864), 
Smith et al., Ecology 2009 (doi:10.1890/08-1815.1).
Actually LAN K. KNAPP and MELINDA D. SMITH did a lot of work on this to access changes in 
plant community response to extreme events 
(this review article: Felton et al. 2017, Phil. Trans. R. Soc. B 372: 20160142 is a nice summary).
-->
<!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
in the appropriate literature, and that is great!-->
through 
life-history traits, alter community structures through cascading effect on 
higher hierarchical levels and lead to local extinction of 
species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
<!--Should I mention idea of match-mismatch hypothesis?-->
For example, a study [@chiu2013impact] showed extreme flood affected 
the annual survival of a prey community that in turn affected predators in 
aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
@frederiksen2008demographic used $43$ years of ring-recovery data 
to examine the demographic impact of ECEs 
through temporal variability in juvenile, immature and adult survival
of European shags on the Isle of May in eastern Scotland.
A relatively recent study [@mcdermott2017sensitivity] investigated the sensitivity 
of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
extreme precipitation, drought) over $37$ years at the site level, across 
each species' life-stages. Experimental study showed successive ECEs with short time intervals 
could weaken the resistance of plant communities too [@dreesen2014successive]. 
All these studies indicated that environmental stochasticity (through ECEs) 
can lead to demographic stochasticity (through influence on reproductive or 
mortality rate) of different age class of a population making it vulnerable 
to local extinction. 
<!--***DAN: Shyamolina, demogra[hic stochasticity is a specific thing. I suspect you may
just want to say instead something like "All these studies indicated that environmental stochasticity (through ECEs) can make a population vulnerable to extirpation."] -->

@@

<!--concept for correlation in the extreme tail: tail-association-->
Though earlier literature considered the influence of spatial and/ or temporal 
correlations in environmental variables on extinction risk of a metapopulation, to 
our knowledge the separate influence on extinction risk of 'tail associations' between
environmental time series in different locations has hardly been considered.

>>>>

***Shyamolina: species reordering is used to access the change in community 
in response to extreme events or global environmental changes, it relates to change 
in relative abundances of dominant species (mainly). For example, as different species have different 
sensitivity levels to extreme condition (means the extreme threshold is not 
constant for all the species in a community), some years with extreme drought favors say c3 grass 
but in the next year c4 grass could be dominant. Such alteration in community composition has "impact on 
ecosystem function, leading to an extreme ecological response". For detailed overview you may see 
@hoover2014resistance, Jones et al. Ecol. lett. 2017 (doi: 10.1111/ele.12864), 
Smith et al., Ecology 2009 (doi:10.1890/08-1815.1).
Actually LAN K. KNAPP and MELINDA D. SMITH did a lot of work on this to access changes in 
plant community response to extreme events 
(this review article: Felton et al. 2017, Phil. Trans. R. Soc. B 372: 20160142 is a nice summary).
-->
<!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
in the appropriate literature, and that is great!-->
through 
life-history traits, alter community structures through cascading effect on 
higher hierarchical levels and lead to local extinction of 
species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
<!--Should I mention idea of match-mismatch hypothesis?-->
For example, a study [@chiu2013impact] showed extreme flood affected 
the annual survival of a prey community that in turn affected predators in 
aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
@frederiksen2008demographic used $43$ years of ring-recovery data 
to examine the demographic impact of ECEs 
through temporal variability in juvenile, immature and adult survival
of European shags on the Isle of May in eastern Scotland.
A relatively recent study [@mcdermott2017sensitivity] investigated the sensitivity 
of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
extreme precipitation, drought) over $37$ years at the site level, across 
each species' life-stages. Experimental study showed successive ECEs with short time intervals 
could weaken the resistance of plant communities too [@dreesen2014successive]. 
All these studies indicated that environmental stochasticity (through ECEs) can 
make a population vulnerable to extirpation.
<!--All these studies indicated that environmental stochasticity (through ECEs) 
can lead to demographic stochasticity (through influence on reproductive or 
mortality rate) of different age class of a population making it vulnerable 
to local extinction.--> 
<!--***DAN: Shyamolina, demogra[hic stochasticity is a specific thing. I suspect you may
just want to say instead something like "All these studies indicated that environmental stochasticity (through ECEs) can make a population vulnerable to extirpation."] -->

@@

<!--concept for correlation in the extreme tail: tail-association-->
Though earlier literature considered the influence of spatial and/ or temporal 
correlations in environmental variables on extinction risk of a metapopulation, to 
our knowledge the separate influence on extinction risk of "tail associations" between
environmental time series in different locations has hardly been considered. "Tail associations" are
the association between two random variables in their extreme values. If two positively correlated 
variables are more associated in their smaller values than larger, they are said to have stronger
"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). Similarly, if two positively correlated 
variables are more associated in their larger values than smaller, they are said to have stronger
"right-tail" or "upper-tail" association (Figure \ref{fig_intro_cop}C, D). Figure \ref{fig_intro_cop} shows 
two random variables can have a wide range of assciation from extreme to moderate degree in their extreme tails, but
with a fixed Kendall's $\tau$.
